PHAT RAID, yo

### 2014-05-12 paper/paper.tex
More updates to paper

@@

@@

@@

@@

Each of these are implemented using Erlang behaviors. An Erlang behavior is a
framework which implements common patterns like servers and finite state
machines. In particular, the supervisor uses the \texttt{supervisor} behavior,
the server and file system both use the \texttt{gen\_server} behavior, and VR
uses the \texttt{gen\_fsm} behavior.

@@

\lstinputlisting[firstline=6,lastline=18,float,
                 caption=Supervisor Behavior --- \texttt{phat.erl},label=supervisor,
                 numbers=left, firstnumber=6]
                {../phat.erl}

Listing \ref{supervisor} is taken from our supervisor code. When a supervisor is
started, the \texttt{supervisor} behavior looks for a unary procedure called
\lstinline!init!. The \lstinline!init! procedure returns a child specification which
is a pair of a restart specification and a list of children. In Listing
\ref{supervisor}, the restart specification, \lstinline!{one_for_all,1,5}!
specifies that if any one node dies, all nodes should be restarted, unless more
than one failure has occurred in the past five seconds. If more than one failure
occurs in five seconds, the supervisor kills all the children and shuts itself
down as well.

@@

@@

>>>>

\usepackage{cite}
% for BibTeX

@@

\usepackage{tikz}
\usetikzlibrary{graphdrawing}

@@

\subsection{System Organization}

@@

\subsection{Erlang and Behaviors}

@@

The four components are implemented using Erlang behaviors. An Erlang behavior
is a framework which implements common patterns like servers and finite state
machines. In particular, the supervisor uses the \texttt{supervisor} behavior,
the server and file system both use the \texttt{gen\_server} behavior, and VR
uses the \texttt{gen\_fsm} behavior.

@@

\subsubsection{The \texttt{supervisor} Behavior}

\lstinputlisting[firstline=6,lastline=18,float,
                 caption=The \texttt{supervisor} Behavior --- \texttt{phat.erl},
                 label=lst:supervisor,
                 numbers=left, firstnumber=6]
                {../phat.erl}

Listing \ref{lst:supervisor} is taken from our supervisor code. When a
supervisor is started, the \texttt{supervisor} behavior looks for a unary
procedure called \lstinline!init!. The \lstinline!init! procedure returns a
child specification which is a pair of a restart specification and a list of
children. In Listing \ref{lst:supervisor}, the restart specification,
\lstinline!{one_for_all,1,5}! specifies that if any one node dies, all nodes
should be restarted, unless more than one failure has occurred in the past five
seconds. If more than one failure occurs in five seconds, the supervisor kills
all the children and shuts itself down as well.

@@

\subsubsection{The \texttt{gen\_fsm} Behavior}

\begin{lstlisting}[float,caption=The \texttt{gen\_fsm} Behavior --- \texttt{vr.erl},
                   label=lst:genfsm, numbers=left, firstnumber=176]
replica( {prepare, MasterViewNumber, _, _, _, _, _}
       , State = #{ viewNumber := ViewNumber})
    when MasterViewNumber > ViewNumber ->
    ?debugFmt("my view is out of date, need to recover~n", []),
    startRecovery(State);

replica( {prepare, MasterViewNumber, _, _, _, _, _}
       , State = #{ timeout := Timeout, viewNumber := ViewNumber})
    when MasterViewNumber < ViewNumber ->
    ?debugFmt("ignoring prepare from old view~n", []),
    {next_state, replica, State, Timeout};

replica( { prepare, _, Op, OpNumber
         , MasterCommitNumber, Client, RequestNum}
       , State = #{ prepareBuffer := PrepareBuffer }) ->
    Message = {OpNumber, Op, Client, RequestNum},
    NewPrepareBuffer = lists:sort([Message|PrepareBuffer]),
    processPrepareOrCommit( OpNumber
                          , MasterCommitNumber
                          , NewPrepareBuffer
                          , State
                          );
\end{lstlisting}

\begin{lstlisting}[float,caption={Processing the VR Message Queue, Part 1 --- \texttt{vr.erl}},
                   label=lst:processing1, numbers=left, firstnumber=375]
processPrepareOrCommit( OpNumber, MasterCommitNumber, NewPrepareBuffer
                      , State = #{ timeout := Timeout
                                 , commitNumber := CommitNumber
                                 , masterNode := MasterNode
                                 , myNode := MyNode
                                 , viewNumber := ViewNumber
                                 , allNodes := Nodes})
  when CommitNumber > MasterCommitNumber ->
    NewViewNumber = ViewNumber + 1,
    NewMaster = chooseMaster(State, NewViewNumber),
    sendToReplicas( MasterNode
                  , Nodes
                  , {startViewChange, NewViewNumber, MyNode}
                  ),
    { next_state
    , viewChange
    , State#{ viewNumber := NewViewNumbern
            , masterNode := NewMaster }
    , Timeout
    };
\end{lstlisting}
\begin{lstlisting}[float,caption={Processing the VR Message Queue, Part 2 --- \texttt{vr.erl}},
                   label=lst:processing2, numbers=left, firstnumber=397]
processPrepareOrCommit( OpNumber, MasterCommitNumber, NewPrepareBuffer
                      , #{ timeout := Timeout } = State) ->
    AfterBuffer =
      processBuffer( State#{ prepareBuffer := NewPrepareBuffer }
                   , NewPrepareBuffer
                   , MasterCommitNumber
                   ),
    AfterLog = processLog(AfterBuffer, MasterCommitNumber),
    #{ masterNode := MasterNode
     , viewNumber := ViewNumber
     , commitNumber := CommitNumber
     , myNode := MyNode } = AfterLog,
    if
        CommitNumber < MasterCommitNumber ->
            startRecovery(State);
        true ->
            sendToMaster( MasterNode
                        , {prepareOk, ViewNumber, OpNumber, MyNode}
                        ),
            {next_state, replica, AfterLog, Timeout}
    end.
\end{lstlisting}

A state in the \texttt{gen\_fsm} behavior manifests as a procedure which accepts
two arguments: the incoming message and the store\footnote{Mutable state is
  implemented as a pure store value which is passed around by the behavior}. The
return value of a state-procedure is usually the 4-tuple
\lstinline!{next_state,NextState,NewStore,Timeout}!. The \texttt{gen\_fsm}
behavior will transition to \lstinline!NextState! and set the store to
\lstinline!NewStore!. When a new message is received the \texttt{gen\_fsm}
behavior will invoke the \lstinline!NextState! procedure with the new message
and the \lstinline!NewStore!.

We implemented Viewstamped Replication\cite{liskov2012viewstamped} to achieve
consensus in the file system. Listing \ref{lst:genfsm} defines how to handle
prepare messages sent to a node in the replica state. It is piecewise defined by
pattern matching on the arguments. The first two definition clauses use the
\lstinline!when! keyword to specify arbitrary, required relationships between
the matched variables.

The final definition clause triggers when the replica receives a message in the
current view. The replica adds the message to a sorted queue and calls a
processing procedure. The processing procedure is defined separately in Listing
\ref{lst:processing1} and Listing \ref{lst:processing2}.

Listing \ref{lst:processing1} handles a message from an out-of-date master,
i.e., his commit number is older than the replica's commit number. The replica
first proposes a view change to the cluster. Afterwards, it changes its
\texttt{gen\_fsm} state to \lstinline!viewChange! and updates the
store\footnote{The store is called \lstinline!State! in the listings} with the
new view number and the new master.

Listing \ref{lst:processing2} handles messages from the current master. The
\lstinline!prepareBuffer! stores prepare messages that arrived out of order. In
particular, if the log ends at operation number $n$ and message $n+1$ is dropped
by the network, all subsequent messages will be buffered. The buffered messages
will not be added to the log until message $n+1$ is received. The call to
\lstinline!processBuffer! moves messages to the log, if all previous messages
have now been received. The call to \lstinline!processLog! commits new messages
in the log which have been committed by the master node. Lines 405 to 408
destructure \lstinline!AfterLog! and bind variables for later use. The final
\lstinline!if! statement responds to the master unless the master has committed
messages we haven't received.\footnote{This can happen if the network drops a message whose
  operation number lies between the replica's last commit number and the
  master's last commit number.}

@@

\bibliographystyle{plain}
\bibliography{paper}
